---
title: "Cameron's Book TS Project"
author: "Cameron Stewart"
date: "11/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Load Packages

```{r}
library(RCurl)
library(tswge)
library(nnfor)
library(vars)
library(tidyverse)
library(lubridate)
library(tseries)
library(RColorBrewer)
```

### Defined Functions

# Bring in peak finding function

```{r}
#Can also be used to find valleys if you put -x in instead of x for realization
find_peaks <- function (x, m = 3){
    shape <- diff(sign(diff(x, na.pad = FALSE)))
    pks <- sapply(which(shape < 0), FUN = function(i){
       z <- i - m + 1
       z <- ifelse(z > 0, z, 1)
       w <- i + m + 1
       w <- ifelse(w < length(x), w, length(x))
       if(all(x[c(z : i, (i + 2) : w)] <= x[i + 1])) return(i + 1) else return(numeric(0))
    })
     pks <- unlist(pks)
     pks
}
```

# Bring in valley finding function

```{r}
find_valleys <- function (x, m = 3){
    x=-x
    shape <- diff(sign(diff(x, na.pad = FALSE)))
    pks <- sapply(which(shape < 0), FUN = function(i){
       z <- i - m + 1
       z <- ifelse(z > 0, z, 1)
       w <- i + m + 1
       w <- ifelse(w < length(x), w, length(x))
       if(all(x[c(z : i, (i + 2) : w)] <= x[i + 1])) return(i + 1) else return(numeric(0))
    })
     pks <- unlist(pks)
     pks
}
```

# Bring in rolling window ASE function

```{r}
Rolling_Window_ASE = function(series, trainingSize, horizon = 1, s = 0, d = 0, phis = 0, thetas = 0)
{
ASEHolder = numeric()

for( i in 1:(length(series)-(trainingSize + horizon) + 1))
{
  
  forecasts = fore.aruma.wge(series[i:(i+(trainingSize-1))],phi = phis, theta = thetas, s = s, d = d,n.ahead = horizon)
  
  ASE = mean((series[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] - forecasts$f)^2)
         
  ASEHolder[i] = ASE

}

ASEHolder
hist(ASEHolder)
WindowedASE = mean(ASEHolder)

print("The Summary Statistics for the Rolling Window ASE Are:")
print(summary(ASEHolder))
print(paste("The Rolling Window ASE is: ",WindowedASE))
return(WindowedASE)
}
```

### Summarized EDA

# Read in the data set data types

```{r}
covid <- read.csv(text=getURL("https://raw.githubusercontent.com/C-Stewart-GH/Time_Series_Project/main/Raw_Data_Files/merged_data.csv"))
covid$Date=mdy(covid$Date)
str(covid)
summary(covid)

cases=read.csv(text=getURL("https://raw.githubusercontent.com/C-Stewart-GH/Time_Series_Project/main/Raw_Data_Files/Texas%20COVID-19%20Case%20Count%20Data%20by%20County.csv"))
cases$Date=mdy(cases$Date)
cases$Case.Count=c(NA,diff(cases$Case.Count))
cases=cases[2:length(cases$Date),]
str(cases)
```

# Plot reduced and full data and check for consistent covariance

```{r}
plotts.sample.wge(covid$case_count,lag.max = 100,trunc = 35)
plotts.sample.wge(cases$Case.Count,lag.max = 40,trunc = 35)

acf(cases$Case.Count[1:(length(cases$Case.Count)/2)],lag.max = 40)
acf(cases$Case.Count[(length(cases$Case.Count)/2+1):length(cases$Case.Count)],lag.max = 40)

acf(covid$case_count[1:(length(covid$case_count)/2)],lag.max = 60)
acf(covid$case_count[(length(covid$case_count)/2+1):length(covid$case_count)],lag.max = 60)
```

# Assume you remove seasonality first

```{r}
fcases_s7=artrans.wge(cases$Case.Count,phi.tr = c(rep(0,6),1),lag.max = 30)
pcases_s7=artrans.wge(covid$case_count,phi.tr = c(rep(0,6),1),lag.max = 60)

#Dicky-Fuller Test shows d=1 does not belong in the data after adding s=7
adf.test(fcases_s7)

#We can see the remaining data is not white noise from the ACF
full_s7=plotts.sample.wge(fcases_s7,lag.max = 50,arlimits = T)

#Use overfit to detect any additional seasonality, none found
factor.wge(phi = c(rep(0,6),1))
est.ar.wge(fcases_s7,p = 17,type = "burg")[0]


#Show there is white noise remaining (p=0 so null is rejected)
ljung.wge(fcases_s7,K = 24)$pval
ljung.wge(fcases_s7,K = 48)$pval


#Use aic5 to find remaining correlation
acf(fcases_s7,lag.max = 60)
pacf(fcases_s7,lag.max = 60)

# aic5.wge(x = fcases_s7,p = 11:20,q = 4:9,type = 'bic')
# aic5.wge(x = fcases_s7,p = 2,q = 9:14,type = 'aic')

#Models based on full data
est_fcases_s7p21q8=est.arma.wge(fcases_s7,p = 21,q=8)

#Run ljung test with autocrrelation
plotts.sample.wge(est_fcases_s7p21q8$res,lag.max = 40,arlimits = T)
acf(est_fcases_s7p21q8$res,lag.max = 50)
pacf(est_fcases_s7p21q8$res,lag.max = 60)
ljung.wge(x = est_fcases_s7p21q8$res,K = 24,p = 2,q = 8)
ljung.wge(x = est_fcases_s7p21q8$res,K = 48,p = 2,q = 8)

#Identify Rolling Window ASE for short term forecast
Rolling_Window_ASE(series = cases$Case.Count,horizon = 7,s = 7,phis = est_fcases_s7p21q8$phi,thetas = est_fcases_s7p21q8$theta,trainingSize = 60)

#Identify Rolling Window ASE for long term forecast
Rolling_Window_ASE(series = cases$Case.Count,horizon = 21,s = 7,phis = est_fcases_s7p21q8$phi,thetas = est_fcases_s7p21q8$theta,trainingSize = 60)
```

# Assume it has no seasonality
```{r}
fcases_d1=artrans.wge(covid$case_count,phi.tr = c(1),lag.max = 40)
pcases_d1=artrans.wge(cases$Case.Count,phi.tr = c(1),lag.max = 40)

#Dicky-Fuller Test shows another d=1 does not belong in the data
adf.test(fcases_d1)

#We can see the remaining data is not white noise from the ACF
full_s7=plotts.sample.wge(fcases_d1,lag.max = 50,arlimits = T)

#Use overfit to detect any additional seasonality, none found
factor.wge(phi = c(rep(0,2),1))
est.ar.wge(fcases_d1,p = 17,type = "burg")[0]

#Show there is white noise remaining (p=0 so null is rejected)
ljung.wge(fcases_d1,K = 24)$pval
ljung.wge(fcases_d1,K = 48)$pval

#Use aic5 to find remaining correlation
acf(fcases_d1,lag.max = 60)
pacf(fcases_d1,lag.max = 60)

#aic5.wge(x = fcases_d1,p = 25:30,q = 10:15,type = 'aic')

#Models based on full data
est_fcases_d1p6q14=est.arma.wge(fcases_d1,p = 6,q=14)

#Run ljung test with autocrrelation
plotts.sample.wge(est_fcases_d1p6q14$res,lag.max = 40,arlimits = T)
acf(est_fcases_d1p6q14$res,lag.max = 60)
pacf(est_fcases_d1p6q14$res,lag.max = 60)
ljung.wge(x = est_fcases_d1p6q14$res,K = 24,p = 2,q = 8)
ljung.wge(x = est_fcases_d1p6q14$res,K = 48,p = 2,q = 8)

#Identify Rolling Window ASE for short term forecast
Rolling_Window_ASE(series = cases$Case.Count,horizon = 7,d = 1,phis = est_fcases_d1p6q14$phi,thetas = est_fcases_d1p6q14$theta,trainingSize = 60)

#Identify Rolling Window ASE for long term forecast
Rolling_Window_ASE(series = cases$Case.Count,horizon = 21,d = 1,phis = est_fcases_d1p6q14$phi,thetas = est_fcases_d1p6q14$theta,trainingSize = 60)
```

# Generate examples from models selected

```{r}
#Original Realization
plotts.wge(cases$Case.Count)

#Generated ARIMA(21,0,8)
plotts.wge(gen.aruma.wge(600,phi = est_fcases_s7p21q8$phi,theta =  est_fcases_s7p21q8$theta,s = 7,vara = est_fcases_s7p21q8$avar,sn = 25300))

#Generated ARIMA(6,1,14)
plotts.wge(gen.aruma.wge(600,phi = est_fcases_d1p6q14$phi,theta =  est_fcases_d1p6q14$theta,d = 1,vara = est_fcases_d1p6q14$avar,sn = 60))
```

### Start Part 2 of Analysis

# Load all data

# Read in the data set
```{r}
covid <- read.csv(text=getURL("https://raw.githubusercontent.com/C-Stewart-GH/Time_Series_Project/main/Raw_Data_Files/merged_data.csv"))
covid$Date=mdy(covid$Date)
str(covid)
summary(covid)
```

# Visualize full data and create mean mobility
```{r}
str(covid)
names(covid)
covid$mobility_mean=rowMeans(x = covid[,c(4:9)],dims = 1)
plotts.sample.wge(covid$case_count)
plotts.sample.wge(covid$tests_taken)
plotts.sample.wge(covid$vaccine_doses_administered)
plotts.sample.wge(covid$retail_and_recreation_percent_change_from_baseline)
plotts.sample.wge(covid$grocery_and_pharmacy_percent_change_from_baseline)
plotts.sample.wge(covid$parks_percent_change_from_baseline)
plotts.sample.wge(covid$transit_stations_percent_change_from_baseline)
plotts.sample.wge(covid$workplaces_percent_change_from_baseline)
plotts.sample.wge(covid$residential_percent_change_from_baseline)
plotts.sample.wge(covid$mobility_mean)
```

# Look at cross correlation of each variable
```{r}
par(mfrow=c(1,1))
covid_d1=artrans.wge(covid$case_count,1)
covid_d1s7=artrans.wge(covid_d1,c(rep(0,6),1))
Ccf(covid_d1,artrans.wge(covid$tests_taken,c(1)),lag.max = 40) #Differenced Lag 4
Ccf(covid_d1,artrans.wge(covid$vaccine_doses_administered,c(1)),lag.max = 40) #Seasonal trend remaining
Ccf(covid_d1,artrans.wge(artrans.wge(covid$vaccine_doses_administered,c(1)),c(rep(0,6),1)),lag.max = 40) #Diff and Seas Lag 19
Ccf(covid_d1,artrans.wge(covid$retail_and_recreation_percent_change_from_baseline,c(rep(0,6),1)),lag.max = 40) #Lag 1
Ccf(covid_d1,artrans.wge(covid$grocery_and_pharmacy_percent_change_from_baseline,c(rep(0,6),1)),lag.max = 40) #Lag 1
Ccf(covid_d1,artrans.wge(covid$parks_percent_change_from_baseline,c(rep(0,6),1)),lag.max = 40) #Insignificant
Ccf(covid_d1,artrans.wge(covid$transit_stations_percent_change_from_baseline,c(rep(0,6),1)),lag.max = 40) #Lag 1
Ccf(covid_d1,artrans.wge(covid$workplaces_percent_change_from_baseline,c(rep(0,6),1)),lag.max = 40) #Lag 0/ Lag 1
Ccf(covid_d1,artrans.wge(covid$residential_percent_change_from_baseline,c(rep(0,6),1)),lag.max = 40) #Lag -1,0,1
Ccf(covid_d1,artrans.wge(covid$mobility_mean,c(rep(0,6),1)),lag.max = 40) #Lag -1
```

# Create VAR Model
```{r}
# #Stationarize all the data
# covid_transformed=data.frame('case_count_d1'=artrans.wge(covid$case_count,1),'tests_taken_d1'=artrans.wge(covid$tests_taken,c(1)),'vax_doses_d1s7'=c(rep(0,7),artrans.wge(artrans.wge(covid$vaccine_doses_administered,c(1)),c(rep(0,6),1),lag.max = 40)),'mob_retail_and_rec_s7'=c(rep(0,6),artrans.wge(covid$retail_and_recreation_percent_change_from_baseline,c(rep(0,6),1))),'mob_grocery_and_pharm_s7'=c(rep(0,6),artrans.wge(covid$grocery_and_pharmacy_percent_change_from_baseline,c(rep(0,6),1))),'mob_transit_stations_s7'=c(rep(0,6),artrans.wge(covid$transit_stations_percent_change_from_baseline,c(rep(0,6),1))),'mob_workplaces_s7'=c(rep(0,6),artrans.wge(covid$workplaces_percent_change_from_baseline,c(rep(0,6),1))),'mob_residential_s7'=c(rep(0,6),artrans.wge(covid$residential_percent_change_from_baseline,c(rep(0,6),1))),'mob_mean_s7'=c(rep(0,6),artrans.wge(covid$mobility_mean,c(rep(0,6),1))))
# 
# head(covid_transformed,20)
# 
# #Slice off the lost rows due to transform
# covid_transformed_sliced=covid_transformed[7:length(covid_transformed$case_count_d1),]
# 
# head(covid_transformed_sliced,20)
# 
# #Use VARSelect to model data
# VARselect(covid_transformed_sliced,lag.max = 50,type = 'const')

#Create first differenced data set and reduce attributes
covid_d1=covid[,2:dim(covid)[2]]
for (i in c(1:(dim(covid)[2]-1))){
  covid_d1[,i]=c(NA,artrans.wge(x = covid_d1[,i],1))
}
names(covid_d1)
str(covid)
str(covid_d1)
covid_d1=covid_d1[2:dim(covid_d1)[1],c(1,2,9,10)]

#Create reduced variable covid data set to mob mean, vax, test count, and case count
names(covid)
covid_reduced=covid[,c(2,3,10,11)]

#Test original data set
VARselect(covid_reduced,lag.max = 45,type = 'const')
  #AIC 8, BIC 8

#Fit based on AIC
fit1a=VAR(covid_reduced,p=8,type="const")
summary(fit1a)
preds=predict(fit1a,n.ahead=21)
par(mfrow=c(1,1))

#Fan charts
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues"))

#Entire Plot
plot(seq(1,dim(covid_reduced)[1],1),covid_reduced$case_count, type = "l")
lines(seq((dim(covid_reduced)[1]-20),dim(covid_reduced)[1],1),preds$fcst$case_count[,1],type = "l",col='blue')

#Visualize only forecasted points
plot(tail(covid_reduced$case_count,21), type = "l")
lines(preds$fcst$case_count[,1],type = "l",col='blue')

short_ASE_fit1a = mean((tail(covid_reduced$case_count,21)[1:7]-preds$fcst$case_count[1:7,1])^2) 
short_ASE_fit1a
short_ASE_fit1a^.5
#7 Day RMSE of 1763.7 which is much lower than the 2981 from our ARIMA(6,1,14) model
#Maybe try to look up rolling window version

long_ASE_fit1a = mean((tail(covid_reduced$case_count,21)-preds$fcst$case_count[,1])^2) 
long_ASE_fit1a
long_ASE_fit1a^.5
#RMSE of 1326.6 which is much lower than the 3806 from our ARIMA(6,1,14) model
#Maybe try to look up rolling window version

#Fit based on BIC
#Same

#Test just a seasonal data set
VARselect(covid_reduced,lag.max = 25,type = 'const',season = 7)
  #AIC 8, BIC 2

#Fit based on AIC
fit2a=VAR(covid_reduced,p=8,type="const",season = 7)
summary(fit2a)
preds=predict(fit2a,n.ahead=21)
par(mfrow=c(1,1))

#Fan charts
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues"))

#Entire Plot
plot(seq(1,dim(covid_reduced)[1],1),covid_reduced$case_count, type = "l")
lines(seq((dim(covid_reduced)[1]-20),dim(covid_reduced)[1],1),preds$fcst$case_count[,1],type = "l",col='blue')

#Visualize only forecasted points
plot(tail(covid_reduced$case_count,21), type = "l",ylim=c(400,5900))
lines(preds$fcst$case_count[,1],type = "l",col='blue')

short_ASE_fit2a = mean((tail(covid_reduced$case_count,21)[1:7]-preds$fcst$case_count[1:7,1])^2) 
short_ASE_fit2a
short_ASE_fit2a^.5
#7 Day RMSE of 2066.2 which is much lower than the 2981 from our ARIMA(6,1,14) model but higher than fit 1a
#Maybe try to look up rolling window version

long_ASE_fit2a = mean((tail(covid_reduced$case_count,21)-preds$fcst$case_count[,1])^2) 
long_ASE_fit2a
long_ASE_fit2a^.5
#21 Day RMSE of 1522.145 which is much lower than the 3806 from our ARIMA(6,1,14) model but higher than fit 1a
#Maybe try to look up rolling window version

#Fit based on BIC
fit2b=VAR(covid_reduced,p=2,type="const",season = 7)
summary(fit2b)
preds=predict(fit2b,n.ahead=21)
par(mfrow=c(1,1))

#Fan charts
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues"))

#Entire Plot
plot(seq(1,dim(covid_reduced)[1],1),covid_reduced$case_count, type = "l")
lines(seq((dim(covid_reduced)[1]-20),dim(covid_reduced)[1],1),preds$fcst$case_count[,1],type = "l",col='blue')

#Visualize only forecasted points
plot(tail(covid_reduced$case_count,21), type = "l",ylim=c(400,5900))
lines(preds$fcst$case_count[,1],type = "l",col='blue')

short_ASE_fit2b = mean((tail(covid_reduced$case_count,21)[1:7]-preds$fcst$case_count[1:7,1])^2) 
short_ASE_fit2b
short_ASE_fit2b^.5
#7 Day RMSE of 1595.7 which is much lower than the 2981 from our ARIMA(6,1,14) model and higher than fit 1a
#Maybe try to look up rolling window version

long_ASE_fit2b = mean((tail(covid_reduced$case_count,21)-preds$fcst$case_count[,1])^2) 
long_ASE_fit2b
long_ASE_fit2b^.5
#21 Day RMSE of 2635.8 which is much lower than the 3806 from our ARIMA(6,1,14) model but higher than fit 1a
#Maybe try to look up rolling window version

#Test a first differenced data set
VARselect(covid_d1,lag.max = 25,type = 'const')
  #AIC 7, BIC 7

#Fit based on AIC
fit3a=VAR(covid_d1,p=7,type="const")
summary(fit3a)
preds=predict(fit3a,n.ahead=21)
par(mfrow=c(1,1))

#Fan charts
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues"))

#Entire Plot
plot(seq(1,dim(covid_d1)[1],1),covid_d1$case_count, type = "l")
lines(seq((dim(covid_d1)[1]-20),dim(covid_d1)[1],1),preds$fcst$case_count[,1],type = "l",col='blue')


#Visualize only forecasted points
plot(tail(covid_d1$case_count,21), type = "l",ylim=c(min(preds$fcst$case_count[,2]),max(preds$fcst$case_count[,3])))
lines(preds$fcst$case_count[,1],type = "l",col='blue')
lines(preds$fcst$case_count[,2],type = "l",col='blue', lty=2)
lines(preds$fcst$case_count[,3],type = "l",col='blue', lty=2)

short_ASE_fit3a = mean((tail(covid_d1$case_count,21)[1:7]-preds$fcst$case_count[1:7,1])^2) 
short_ASE_fit3a
short_ASE_fit3a^.5
#7 Day RMSE of 1278.8 which is second lowest short term RMSE
#Maybe try to look up rolling window version

long_ASE_fit3a = mean((tail(covid_d1$case_count,21)-preds$fcst$case_count[,1])^2) 
long_ASE_fit3a
long_ASE_fit3a^.5
#21 Day RMSE of 974.3 which is much lower than the 3806 from our ARIMA(6,1,14) model and lowest long term VAR model
#Maybe try to look up rolling window version

#Test a first differenced and seasonal data set
VARselect(covid_d1,lag.max = 25,type = 'const',season = 7)
  #AIC 7, BIC 7

#Fit based on AIC
fit4a=VAR(covid_d1,p=7,type="const",season = 7)
summary(fit4a)
preds=predict(fit4a,n.ahead=21)
par(mfrow=c(1,1))

#Fan charts
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues"))

#Entire Plot
plot(seq(1,dim(covid_d1)[1],1),covid_d1$case_count, type = "l")
lines(seq((dim(covid_d1)[1]-20),dim(covid_d1)[1],1),preds$fcst$case_count[,1],type = "l",col='blue')

#Visualize only forecasted points
plot(tail(covid_d1$case_count,21), type = "l",ylim=c(-3900,4600))
lines(preds$fcst$case_count[,1],type = "l",col='blue')

short_ASE_fit4a = mean((tail(covid_d1$case_count,21)[1:7]-preds$fcst$case_count[1:7,1])^2) 
short_ASE_fit4a
short_ASE_fit4a^.5
#7 Day RMSE of 1143.2 which is lowest short term RMSE
#Maybe try to look up rolling window version

long_ASE_fit4a = mean((tail(covid_d1$case_count,21)-preds$fcst$case_count[,1])^2) 
long_ASE_fit4a
long_ASE_fit4a^.5
#21 Day RMSE of 1185.6 which is much lower than the 3806 from our ARIMA(6,1,14) model and second lowest model
#Maybe try to look up rolling window version

#Create first different response only data set
names(covid_reduced)
cases_d1=artrans.wge(covid_reduced$case_count,1)
covid_reduced_cases_d1=covid_reduced
covid_reduced_cases_d1$case_count_d1=c(NA,cases_d1)
names(covid_reduced_cases_d1)
covid_reduced_cases_d1=covid_reduced_cases_d1[2:dim(covid_reduced_cases_d1)[1],c(1,3,4,5)]
str(covid_reduced_cases_d1)

#Test a first differenced cases only without seasonality
VARselect(covid_reduced_cases_d1,lag.max = 25,type = 'const')
  #AIC 8, BIC 8

#Fit based on AIC
fit5a=VAR(covid_reduced_cases_d1,p=8,type="const")
summary(fit5a)
preds=predict(fit5a,n.ahead=21)
par(mfrow=c(1,1))

#Fan charts
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues"))

#Entire Plot
plot(seq(1,dim(covid_reduced_cases_d1)[1],1),covid_reduced_cases_d1$case_count, type = "l")
lines(seq((dim(covid_reduced_cases_d1)[1]-20),dim(covid_reduced_cases_d1)[1],1),preds$fcst$case_count[,1],type = "l",col='blue')

#Visualize only forecasted points
plot(tail(covid_reduced_cases_d1$case_count,21), type = "l",ylim=c(-3900,4600))
lines(preds$fcst$case_count[,1],type = "l",col='blue')

short_ASE_fit5a = mean((tail(covid_reduced_cases_d1$case_count,21)[1:7]-preds$fcst$case_count[1:7,1])^2) 
short_ASE_fit5a
short_ASE_fit5a^.5
#7 Day RMSE of 1300.18
#Maybe try to look up rolling window version

long_ASE_fit5a = mean((tail(covid_reduced_cases_d1$case_count,21)-preds$fcst$case_count[,1])^2) 
long_ASE_fit5a
long_ASE_fit5a^.5
#21 Day RMSE of 994.35 which is much lower than the 3806 from our ARIMA(6,1,14) model and lowest RMSE

#Test a first differenced cases only with seasonality
VARselect(covid_reduced_cases_d1,lag.max = 25,type = 'const',season = 7)
  #AIC 8, BIC 8

#Fit based on AIC
fit6a=VAR(covid_reduced_cases_d1,p=8,type="const",season = 7)
summary(fit6a)
preds=predict(fit6a,n.ahead=21)
par(mfrow=c(1,1))

#Fan charts
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues"))

#Entire Plot
plot(seq(1,dim(covid_reduced_cases_d1)[1],1),covid_reduced_cases_d1$case_count, type = "l")
lines(seq((dim(covid_reduced_cases_d1)[1]-20),dim(covid_reduced_cases_d1)[1],1),preds$fcst$case_count[,1],type = "l",col='blue')

#Visualize only forecasted points
plot(tail(covid_reduced_cases_d1$case_count,21), type = "l",ylim=c(min(preds$fcst$case_count[,2]),12100))
lines(preds$fcst$case_count[,1],type = "l",col='blue')
lines(preds$fcst$case_count[,2],type = "l",col='blue',lty=2)
lines(preds$fcst$case_count[,3],type = "l",col='blue',lty=2)

short_ASE_fit6a = mean((tail(covid_reduced_cases_d1$case_count,21)[1:7]-preds$fcst$case_count[1:7,1])^2) 
short_ASE_fit6a
short_ASE_fit6a^.5
#7 Day RMSE of 1135.7
#Maybe try to look up rolling window version

long_ASE_fit6a = mean((tail(covid_reduced_cases_d1$case_count,21)-preds$fcst$case_count[,1])^2) 
long_ASE_fit6a
long_ASE_fit6a^.5
#21 Day RMSE of 1116.739 which is much lower than the 3806 from our ARIMA(6,1,14) model and lowest RMSE
```

# Summaryize above results
```{r}
#Minimize Short 7 Day Forecast RMSE selects model fit6a (diff response with seasonality and lag of 8) - RMSE of 1135.75
(short_ASE_fit1a)^.5
(short_ASE_fit2a)^.5
(short_ASE_fit2b)^.5
(short_ASE_fit3a)^.5
(short_ASE_fit4a)^.5
(short_ASE_fit5a)^.5
(short_ASE_fit6a)^.5
min(short_ASE_fit1a,short_ASE_fit2a,short_ASE_fit2b,short_ASE_fit3a,short_ASE_fit4a,short_ASE_fit5a,short_ASE_fit6a)^.5

#Minimize Long 21 Day Forecast RMSE selects model fit3a (diff all data without seasonality lag of 8) - RMSE of 974.29
(long_ASE_fit1a)^.5
(long_ASE_fit2a)^.5
(long_ASE_fit2b)^.5
(long_ASE_fit3a)^.5
(long_ASE_fit4a)^.5
(long_ASE_fit5a)^.5
(long_ASE_fit6a)^.5
min(long_ASE_fit1a,long_ASE_fit2a,long_ASE_fit2b,long_ASE_fit3a,long_ASE_fit4a,long_ASE_fit5a,long_ASE_fit6a)^.5

#Minimize Short 7 Day Forecast AIC selects model fit6a (diff response with seasonality and lag of 8) - AIC of 7659
AIC(fit1a$varresult$case_count)
AIC(fit2a$varresult$case_count)
AIC(fit2b$varresult$case_count)
AIC(fit3a$varresult$case_count)
AIC(fit4a$varresult$case_count)
AIC(fit5a$varresult$case_count)
AIC(fit6a$varresult$case_count)

#Minimize Short 7 Day Forecast BIC selects model fit6a (diff response with seasonality and lag of 8) - BIC of 7819
BIC(fit1a$varresult$case_count)
BIC(fit2a$varresult$case_count)
BIC(fit2b$varresult$case_count)
BIC(fit3a$varresult$case_count)
BIC(fit4a$varresult$case_count)
BIC(fit5a$varresult$case_count)
BIC(fit6a$varresult$case_count)
```

# Create MLP model
```{r}
###MLP with reduced original data

#Create train and whole ts set
covid_reduced_train=covid_reduced[1:(dim(covid_reduced)[1]-21),]

covid_reduced_train$case_count <- ts(covid_reduced_train$case_count, start = decimal_date(as.Date("2020-09-14")), frequency = 365)
covid_reduced_train$tests_taken <- ts(covid_reduced_train$tests_taken, start = decimal_date(as.Date("2020-09-14")), frequency = 365)
covid_reduced_train$vaccine_doses_administered <- ts(covid_reduced_train$vaccine_doses_administered, start = decimal_date(as.Date("2020-09-14")), frequency = 365)
covid_reduced_train$mobility_mean <- ts(covid_reduced_train$mobility_mean, start = decimal_date(as.Date("2020-09-14")), frequency = 365)

covid_reduced_ts=covid_reduced

covid_reduced_ts$case_count <- ts(covid_reduced_ts$case_count, start = decimal_date(as.Date("2020-09-14")), frequency = 365)
covid_reduced_ts$tests_taken <- ts(covid_reduced_ts$tests_taken, start = decimal_date(as.Date("2020-09-14")), frequency = 365)
covid_reduced_ts$vaccine_doses_administered <- ts(covid_reduced_ts$vaccine_doses_administered, start = decimal_date(as.Date("2020-09-14")), frequency = 365)
covid_reduced_ts$mobility_mean <- ts(covid_reduced_ts$mobility_mean, start = decimal_date(as.Date("2020-09-14")), frequency = 365)

#Fit model and forecast
set.seed(2)
fit_mlp1 = mlp(y = covid_reduced_train$case_count,xreg = covid_reduced_train[,c(1,3,4)],hd.auto.type = 'cv',reps =  30,comb = 'median',allow.det.season = T)
fit_mlp1
plot(fit_mlp1)

short_f_mlp1 = forecast(fit_mlp1, h = 7, xreg = covid_reduced_ts[,c(1,3,4)])
plot(short_f_mlp1,xlim=c(2021.7,2022))
plot(seq(1,7),tail(covid_reduced_ts$case_count,21)[1:7],type = "l",ylim=c(-4800,12000)) 
lines(seq(1,7),short_f_mlp1$mean, col = "blue",type= 'l')
short_ASE_fit_mlp1 = mean((tail(covid_reduced_ts$case_count,21)[1:7]-short_f_mlp1$mean)^2) 
short_ASE_fit_mlp1^.5
#RMSE of 3698

long_f_mlp1 = forecast(fit_mlp1, h = 21, xreg = covid_reduced_ts[,c(1,3,4)],level=c(95))

plot(long_f_mlp1,xlim=c(2021.7,2022))


plot(long_f_mlp1$mean,type='l',ylim=c(-20000,20000),lwd=5)
for (i in 1:20){
    lines(long_f_mlp1$all.mean[,i],col = 'grey',type = 'l')
}
lines(tail(covid_reduced_ts$case_count,21),type = "l",col='blue',lwd=5)


plot(seq(1,21),tail(covid_reduced_ts$case_count,21),type = "l",ylim=c(-6500,12000)) 
lines(seq(1,21),long_f_mlp1$mean, col = "blue",type= 'l')
long_ASE_fit_mlp1 = mean((tail(covid_reduced_ts$case_count,21)-long_f_mlp1$mean)^2) 
long_ASE_fit_mlp1^.5
#RMSE 4795

###Use differenced repsonse set

#Create train and whole ts set
covid_reduced_cases_d1_train=covid_reduced_cases_d1[1:(dim(covid_reduced_cases_d1)[1]-21),]

covid_reduced_cases_d1_train$case_count_d1 <- ts(covid_reduced_cases_d1_train$case_count_d1, start = decimal_date(as.Date("2020-09-15")), frequency = 365)
covid_reduced_cases_d1_train$tests_taken <- ts(covid_reduced_cases_d1_train$tests_taken, start = decimal_date(as.Date("2020-09-15")), frequency = 365)
covid_reduced_cases_d1_train$vaccine_doses_administered <- ts(covid_reduced_cases_d1_train$vaccine_doses_administered, start = decimal_date(as.Date("2020-09-15")), frequency = 365)
covid_reduced_cases_d1_train$mobility_mean <- ts(covid_reduced_cases_d1_train$mobility_mean, start = decimal_date(as.Date("2020-09-15")), frequency = 365)

covid_reduced_cases_d1_ts=covid_reduced_cases_d1

covid_reduced_cases_d1_ts$case_count_d1 <- ts(covid_reduced_cases_d1_ts$case_count_d1, start = decimal_date(as.Date("2020-09-15")), frequency = 365)
covid_reduced_cases_d1_ts$tests_taken <- ts(covid_reduced_cases_d1_ts$tests_taken, start = decimal_date(as.Date("2020-09-15")), frequency = 365)
covid_reduced_cases_d1_ts$vaccine_doses_administered <- ts(covid_reduced_cases_d1_ts$vaccine_doses_administered, start = decimal_date(as.Date("2020-09-15")), frequency = 365)
covid_reduced_cases_d1_ts$mobility_mean <- ts(covid_reduced_cases_d1_ts$mobility_mean, start = decimal_date(as.Date("2020-09-15")), frequency = 365)

#Fit model and forecast
set.seed(2)
fit_mlp2 = mlp(y = covid_reduced_cases_d1_train$case_count_d1,xreg = covid_reduced_cases_d1_train[,c(1,2,3)],hd.auto.type = 'cv',reps =  30,comb = 'median',allow.det.season = T)
fit_mlp2
plot(fit_mlp2)

short_f_mlp2 = forecast(fit_mlp2, h = 7, xreg = covid_reduced_cases_d1_ts[,c(1,2,3)])
plot(short_f_mlp2,xlim=c(2021.7,2022))
plot(seq(1,7),tail(covid_reduced_cases_d1_ts$case_count_d1,21)[1:7],type = "l",ylim=c(-4800,12000)) 
lines(seq(1,7),short_f_mlp2$mean, col = "blue",type= 'l')
short_ASE_fit_mlp2 = mean((tail(covid_reduced_cases_d1_ts$case_count_d1,21)[1:7]-short_f_mlp2$mean)^2) 
short_ASE_fit_mlp2^.5
#RMSE of 2760

long_f_mlp2 = forecast(fit_mlp2, h = 21, xreg = covid_reduced_cases_d1_ts[,c(1,2,3)])
plot(long_f_mlp2,xlim=c(2021.7,2022))
plot(seq(1,21),tail(covid_reduced_cases_d1_ts$case_count_d1,21),type = "l",ylim=c(-6500,12000)) 
lines(seq(1,21),long_f_mlp2$mean, col = "blue",type= 'l')
long_ASE_fit_mlp2 = mean((tail(covid_reduced_cases_d1_ts$case_count_d1,21)-long_f_mlp2$mean)^2) 
long_ASE_fit_mlp2^.5
#RMSE 5788
```

# MLP RMSE Summary
```{r}
(short_ASE_fit_mlp1)^.5
(short_ASE_fit_mlp2)^.5

(long_ASE_fit_mlp1)^.5
(long_ASE_fit_mlp2)^.5
```

# Selected MLP Models and Ensemble Model
```{r}
#Selected VAR Model
#Fit based on AIC
fit6a=VAR(covid_reduced_cases_d1,p=8,type="const",season = 7)
summary(fit6a)
preds_6a=predict(fit6a,n.ahead=21)
par(mfrow=c(1,1))

#Fan charts
fanchart(preds_6a, colors = brewer.pal(n = 8, name = "Blues"))

#Entire Plot
plot(seq(1,dim(covid_reduced_cases_d1)[1],1),covid_reduced_cases_d1$case_count, type = "l")
lines(seq((dim(covid_reduced_cases_d1)[1]-20),dim(covid_reduced_cases_d1)[1],1),preds_6a$fcst$case_count[,1],type = "l",col='blue')

#Visualize only forecasted points
plot(tail(covid_reduced_cases_d1$case_count,21), type = "l",ylim=c(min(preds_6a$fcst$case_count[,2]),12100))
lines(preds_6a$fcst$case_count[,1],type = "l",col='blue')
lines(preds_6a$fcst$case_count[,2],type = "l",col='blue',lty=2)
lines(preds_6a$fcst$case_count[,3],type = "l",col='blue',lty=2)

short_ASE_fit6a = mean((tail(covid_reduced_cases_d1$case_count,21)[1:7]-preds_6a$fcst$case_count[1:7,1])^2) 
short_ASE_fit6a
short_ASE_fit6a^.5
#7 Day RMSE of 1135.7
#Maybe try to look up rolling window version

long_ASE_fit6a = mean((tail(covid_reduced_cases_d1$case_count,21)-preds_6a$fcst$case_count[,1])^2) 
long_ASE_fit6a
long_ASE_fit6a^.5
#21 Day RMSE of 1116.739 which is much lower than the 3806 from our ARIMA(6,1,14) model and lowest RMSE



#Selected MLP Model1
covidData = ts(covid$case_count[1:395])
covidXX = data.frame(testsTS = ts(covid$tests_taken[1:395]), vaccinesTS = ts(covid$vaccine_doses_administered[1:395]), park_mobility= ts(covid$parks_percent_change_from_baseline[1:395]))
covidXX_full = data.frame(testsTS = ts(covid$tests_taken), vaccinesTS = ts(covid$vaccine_doses_administered), park_mobility= ts(covid$parks_percent_change_from_baseline))
set.seed(2)
fitCOVIDXX = mlp(covidData,xreg = covidXX)
fcstCOVIDXX= forecast(fitCOVIDXX,h=21,xreg = covidXX_full)

plot(seq(1,21),fcstCOVIDXX$mean,type='l',ylim=c(-20000,20000),lwd=5)
for (i in 1:20){
    lines(seq(1,21),fcstCOVIDXX$all.mean[,i],col = 'grey',type = 'l')
}
lines(tail(covid$case_count,21),type = "l",col='blue',lwd=5)

plot(seq(1,21),fcstCOVIDXX$mean,type='l',ylim=c(-20000,20000),lwd=5)
lines(seq(1,21),fcstCOVID$mean,col = "blue")
lines(seq(1,21),fcstCOVIDXX$mean,col = "red")
lines(seq(1,21), zz, type = "l", col = "green")

ASE_DEEP_LONG_XX = mean((tail(covid$case_count,21) - fcstCOVIDXX$mean)^2)
ASE_DEEP_LONG_XX
RMSE_DEEP_LONG_XX = sqrt(ASE_DEEP_LONG_XX)
RMSE_DEEP_LONG_XX

#Selected MLP Model2
covidData = ts(covid$case_count[1:395])
covidXX2 = data.frame(testsTS = ts(covid$tests_taken[1:395]), vaccinesTS = ts(covid$vaccine_doses_administered[1:395]))
covidXX_full2 = data.frame(testsTS = ts(covid$tests_taken), vaccinesTS = ts(covid$vaccine_doses_administered))
set.seed(2)
fitCOVIDXX2 = mlp(covidData,xreg = covidXX2)
fcstCOVIDXX2= forecast(fitCOVIDXX2,h=21,xreg = covidXX_full2)

plot(seq(1,21),fcstCOVIDXX2$mean,type='l',ylim=c(-20000,20000),lwd=5)
for (i in 1:20){
    lines(seq(1,21),fcstCOVIDXX2$all.mean[,i],col = 'grey',type = 'l')
}
lines(tail(covid$case_count,21),type = "l",col='blue',lwd=5)

plot(seq(1,21),fcstCOVIDXX2$mean,type='l',ylim=c(-20000,20000),lwd=5)
lines(seq(1,21),fcstCOVID$mean,col = "blue")
lines(seq(1,21),fcstCOVIDXX2$mean,col = "red")
lines(seq(1,21), zz, type = "l", col = "green")

ASE_DEEP_LONG_XX2 = mean((tail(covid$case_count,21) - fcstCOVIDXX2$mean)^2)
ASE_DEEP_LONG_XX2
RMSE_DEEP_LONG_XX2 = sqrt(ASE_DEEP_LONG_XX)
RMSE_DEEP_LONG_XX2

#Ensemble forecast
adj_preds2=preds_6a$fcst$case_count_d1[,1]+tail(covid$case_count,22)[1:21]
plot(adj_preds2,type='l',ylim=c(-10000,10000))
lines(tail(covid$case_count,21),col='blue')

adj_ASE_6a = mean((tail(covid$case_count,21)-adj_preds2)^2) 
adj_ASE_6a^.5

ensemble_preds=(adj_preds2+fcstCOVIDXX$mean)/2
plot(tail(covid$case_count,21),type='l',ylim=c(-10000,10000))
lines(seq(1,21),tail(ensemble_preds,21),col='blue')

ensemble_ASE = mean((tail(covid$case_count,21)[1:7]-ensemble_preds[1:7])^2) 
ensemble_ASE^.5

ensemble_ASE = mean((tail(covid$case_count,21)-ensemble_preds)^2) 
ensemble_ASE^.5

#Final comparison plot
plot(tail(covid$case_count,21),type='l',ylim=c(0,7000),lwd=4)
lines(adj_preds2,type = "l",col='grey')
lines(seq(1,21),fcstCOVIDXX$mean,col = "grey")
lines(seq(1,21),tail(ensemble_preds,21),col='blue',lwd=2,lty=2)

#Final comparison plot
plot(tail(covid$case_count,100),type='l',ylim=c(0,30000),lwd=2)
lines(seq(80,100),adj_preds2,type = "l",col='dark grey')
lines(seq(80,100),fcstCOVIDXX$mean,col = "dark grey")
lines(seq(80,100),tail(ensemble_preds,21),col='blue',lwd=2,lty=2)


```

